High geomagnetic field intensity recorded by anorthosite xenoliths requires a strongly powered late Mesoproterozoic geodynamo

### 2022-02-09 PNAS_manuscript/PNAS_manuscript.tex

\begin{figure*}[h!]
\noindent\includegraphics[width=17.8 cm]{Paleointensity_plot_cooling_corrected.pdf}
\centering
\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site (black bars with grey uncertainty boxes) and locality level (orange dashed line and bar). All results are corrected for cooling rate with a factor of 0.74. }}
\label{fig:PINT_cooling_corrected}
\end{figure*}

@@

\begin{figure*}[h!]
\noindent\includegraphics[width=17.8 cm]{PINT_compilation.pdf}
\centering
\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}), including all Phanerozoic VDM and VADM records with Q$_{PI}$ values $>$3 and additional Neoproterozoic data from refs. \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}. Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text. Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic. The highest site-mean virtual dipole moment of the anorthosites would rank in the top 1$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core. }}
\label{fig:PINT_compilation}
\end{figure*}

>>>>

\begin{figure*}[h!]
\noindent\includegraphics[width=17.8 cm]{Paleointensity_plot_cooling_corrected.pdf}
\centering
\caption{\small{Summary plot of individual specimen absolute paleointensity results (square symbols) from this study and their averages and standard deviations at site (black bars with grey uncertainty boxes) and locality level (orange dashed line and bar). All results are corrected for cooling rate with a factor of 0.74. As discussed in the text, the Beaver River anorthosite xenoliths likely record geomagnetic field intensities at similar but distinct times during the cooling of the Beaver River diabase. To compare these observations with more recent paleointensity variability, we took random draws of 3 values from the PADM2M model \cite{Ziegler2011a} for the VADM during the past 2 million years for a total of 10000 iterations and show the distribution of the calculated paleointensity average value of each draw at a latitude of 22\textdegree N. }}
\label{fig:PINT_cooling_corrected}
\end{figure*}

@@

\begin{figure*}[h!]
\noindent\includegraphics[width=17.8 cm]{PINT_compilation.pdf}
\centering
\caption{\footnotesize{Compilation of calculated virtual (axial) dipole moment values from the PINT database (PINT v8.0.0; \url{http://www.pintdb.org/};  \citealp{Bono2021a}), including all Phanerozoic VDM and VADM records with Q$_{PI}$ values $>$3 and additional Neoproterozoic data from refs. \citealp{Lloyd2021a, Lloyd2021b, Thallner2021a, Thallner2021b}. Paleointensity estimates from \cite{Pesonen1983a} and \cite{Kulakov2013a} are not included in the compilation due to the specimen-level double-slope behavior as discussed in the text. Overall the anorthosite xenoliths from this study record a high Mesoproterozoic field exceeding the value projected by the second order polynomial curve from \cite{Bono2019a} which is based on an interpretation of there being a monotonic decay of the geodynamo through the Proterozoic. The highest site-mean virtual dipole moment of the anorthosites would rank in the top 2$\%$ of those in the database for the Cenozoic Era (the last 66 million years) when there was unequivocally a crystallizing inner core. The y axis range is set such that the top 1\% of the Cenozoic paleointensity data is cut off.}}
\label{fig:PINT_compilation}
\end{figure*}
